Awareness and Russell's Epistemology

### 2011-01-27 foundationalism.tex
Fixed reference to Fagin et al, reworded motivation of the formalism

\subsection{Modeling Russell Using Awareness Logic}

In this section we introduce models of Russell's epistemology.  We
will make use of \emph{awareness models} $\mathbb{M}$, using semantics originally developed in \cite{fagin_belief_198}.

@@

All that is left is to establish a connection between knowledge by
description and logical
consequence.  Let $\vdash$ be any Hilbert-style calculus,
that is a sequent that takes a set of premises $\Gamma$ on the left and yields a
single conclusion $\phi$ on the right. Define the following terminology:
\begin{itemizedot}
\item $\vdash$ is \emph{relatively sound} for $\mathbb{M}$ if and only if
  when $\Gamma \vdash \phi$, then if $\mathbb{M},w\models \Gamma$ then
  $\mathbb{M},w \models \phi$.
\item $\vdash$ is \emph{reflexive} if and only if whenever $\phi \in
  \Gamma$ then $\Gamma \vdash \phi$
\item $\vdash$ makes true \emph{modus ponens} if and only if whenever
  $\Gamma \vdash \phi \to \psi$ and $\Gamma \vdash \phi$, then $\Gamma
  \vdash \psi$
\end{itemizedot}

A model $\mathbb{M}$ is thought to be suitable for Russell's epistemology just in
case there is some relatively sound,
reflexive Hilbert-style calculus making true modus ponens, such that for all
worlds $w$:

@@

Here $Th(\mathbb{M})$ is the \emph{theory of $\mathbb{M}$}, and
denotes the set $\{ \phi\ |\ \mathbb{M},w \models \phi$ for all worlds
$w \}$.  We take it to represent Russell's \emph{intuitive knowledge}
in our modeling\emph{knowledge by description}\footnote{A caveat: we are being very liberal with our interpretation
  of Russell's
  epistemology here.  The set $Th(\mathbb{M})$ contains, among other
  things, all tautologies for any logic $\mathbb{M}$ obeys.  Russell
  intended for intuitive knowledge to correspond to empirical
  knowledge and a small number of a priori truths, which serve as
  an axiom system for derivations.  We defend choice to call $Th(\mathbb{M})$ intuitive
  knowledge, since (1) it may be considered as a sort of a priori knowledge, (2) it fits the definition of knowledge by description well,
  and (3) it provides a framework for our subsequent formal results.}.

@@

Restricting our attention to models where implicit knowledge is
equated to the presence
of a derivation is a faithful to Russell's epistemology.  Our response
to Hintikka's criticism of this is that while epistemic logic may be
based on models, it could also be based on particular, special models
which have an associated logic suited to them.  Assuming that
knowledge in epistemic logic corresponds to \emph{some} notion of logical consequence
does not rely on the notion of logical consequence in epistemic
itself.  While logic about logic is circular, it is not
\emph{viciously} so.

The rest of this essay is devoted to developing awareness logics which have
the mechanics we have sketched here.

>>>>

\subsection{Modeling Russell}\label{mod-russell}

In this section we introduce models of Russell's epistemology.  We
will make use of \emph{awareness models} $\mathbb{M}$, using semantics originally developed in \cite{fagin_belief_1988}.

@@

All that is left is to establish a connection between knowledge by
description and logical
consequence.  Let $\vdash$ be any Hilbert-style calculus,
that is a sequent that takes a set of premises $\Gamma$ on the left and yields a
single conclusion $\phi$ on the right. Define the following terminology:
\begin{itemizedot}
\item $\vdash$ is \emph{relatively sound} for $\mathbb{M}$ if and only if
  when $\Gamma \vdash \phi$, then if $\mathbb{M},w\models \gamma$, for
  all $\gamma \in \Gamma$, then
  $\mathbb{M},w \models \phi$.
\item $\vdash$ is \emph{reflexive} if and only if whenever $\phi \in
  \Gamma$ then $\Gamma \vdash \phi$
\item $\vdash$ makes true \emph{modus ponens} if and only if whenever
  $\Gamma \vdash \phi \to \psi$ and $\Gamma \vdash \phi$, then $\Gamma
  \vdash \psi$
\item $\vdash$ is said to be \emph{suitable} for $\mathbb{M}$ just in
  case that it is reflexive, makes true modus ponens, and is
  relatively sound for $\mathbb{M}$.
\end{itemizedot}

A model $\mathbb{M}$ is thought to model Russell's epistemology in the
case that there is some suitable logic $\vdash$ such that for all
worlds $w$:

@@

Here $Th(\mathbb{M})$ is the \emph{theory of $\mathbb{M}$}, and
denotes the set $\{ \phi\ |\ \mathbb{M},w \models \phi$ for all worlds
$w \}$.  We take it to represent Russell's \emph{intuitive
  knowledge} \footnote{A caveat: we are being very liberal with our interpretation
  of Russell's
  epistemology here.  The set $Th(\mathbb{M})$ contains, among other
  things, all tautologies for any logic $\mathbb{M}$ obeys.  Russell
  intended for intuitive knowledge to correspond to empirical
  knowledge and a small number of a priori truths, which serve as
  an axiom system for derivations.  We defend choice to call $Th(\mathbb{M})$ intuitive
  knowledge, since (1) it may be considered as a sort of a priori knowledge, (2) it fits the definition of knowledge by description well,
  and (3) it provides a framework for our subsequent formal
  results.}.

@@

Models where implicit knowledge is
equated to the presence
of a derivation are faithful to Russell's epistemology.  As we stated
before, Russell's knowledge by description is the logical closure of
acquaintance and intuitive knowledge, which is exactly our proposal.  Our response
to Hintikka's criticism of this approach is that while epistemic logic may be
based on models, it could also be based on particular models
which have logic associated with them in a special way.  Assuming that
knowledge in epistemic logic corresponds to \emph{some} notion of logical consequence
does not require us to have already formulated the notion of logical
consequence we will ultimately
develop.  While logic about logic is circular, it is not
\emph{viciously} so, and moreover poses no inescapable paradoxes.

The rest of this essay is devoted to developing awareness logics which have
the mechanics we have sketched.
